---
title: "TidyEducation"
output:
 html_document:
   toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load `tidyverse` package

```{r}
library(tidyverse)
```

# Classifying data types

Before identifying the data types, the dataset will be reduced to remove irrelevant columns.

```{r}
totaledu <- read.csv("./edu.csv")
edu <- totaledu %>% select(
  EstablishmentName,
  EstablishmentTypeGroup..code.,
  Boarders..code.,
  NumberOfPupils,
  NumberOfBoys,
  NumberOfGirls,
  PercentageFSM,
  Town,
  Gender..code.,
  PercentageFSM
)

head(edu)
```

* **Nominal:** TypeOfEstablishment, Boarders, Town, Gender.
  * These variables only support equality operators.
* **Ratio:** NumberOfPupils, NumberOfBoys, NumberOfGirls, PercentageFSM
  * These variables support all mathematical operations.

# Exploratoration

It is of interest to delineate the correlation between different predictors.
I use the term 'predictors' hesitantly, as no hypothesis has currently been set.

```{r}
fsm <- edu %>% arrange(desc(PercentageFSM))

head(fsm)
```

Some statistical summaries will also be included for ratio data types.

```{r}
summary <- edu %>% summarise(
  meanpupils = mean(NumberOfPupils, na.rm = TRUE),
  meanfsm = mean(PercentageFSM, na.rm = TRUE)
)

summary
```
